---
title: "Electricity Demand Forecasting - Linear Difference Eq"
author: "Arunav Saikia"
output: github_document
---
<h3> Load Data </h3>
```{r}
library(fpp2)
plot(elecdaily)
data <- data.frame(elecdaily)
# data <- data.frame(elecdemand)
Y <- data$Demand - mean(data$Demand)
demand <- Y
x <- data$Temperature - mean(data$Temperature)
```
<h3> Fit Models </h3>
Simple Model - $y(t) = b_0 + b_1 · y(t − 1) + b_2 · x(t) + b_3 · cos(2\pi f_1t) + b_4 · sin(2\pi f_1t) + b_5 · cos(2\pi f_2t) + b_6 · sin(2\pi f_2t) + w(t)$
Complex Model - $$y(t) = b_0 + b_1 · y(t − 1)  + b_2 · y(t - 2) + b_3 · y(t - 3) + b_4 · x(t) + b_5 · x(t − 1) + b_6 · x(t - 2) + b_7 · x(t - 3)\\ + b_8 · cos(2\pi f_1t) + b_9 · sin(2\pi f_1t) + b_{10} · cos(2\pi f_2t) + b_{11} ·  sin(2\pi f_2t) + w(t)$$
Third Model -$$y(t) = b_0 + b_1 · y(t − 1) + b_2 · x(t) + b_3 · x(t − 1) + b_4 · x(t - 2)\\ + b_5 · cos(2\pi f_1t) + b_6 · cos(2\pi f_2t) + b_7 · sin(2\pi f_2t) + w(t)$$
```{r}
N = length(Y)

f1 = .1428571 # in previous assignment we found that f1 and f2
f2 = .2857143 # are two frequencies where the signal has seasonal trend
time = 1:N
cos1 = cos(2*pi*time*f1)[4:N]
sin1 = sin(2*pi*time*f1)[4:N]
cos2 = cos(2*pi*time*f2)[4:N]
sin2 = sin(2*pi*time*f2)[4:N]

dY_1 = filter(Y,c(0,1),sides=1)
dY_2 = filter(Y,c(0,0,1),sides=1)
dY_3 = filter(Y,c(0,0,0,1),sides=1)
dx_1 = filter(x,c(0,1),sides=1)
dx_2 = filter(x,c(0,0,1),sides=1)
dx_3 = filter(x,c(0,0,0,1),sides=1)

Y = Y[4:N]
x = x[4:N]
dY_1 = dY_1[4:N]
dY_2 = dY_2[4:N]
dY_3 = dY_3[4:N]
dx_1 = dx_1[4:N]
dx_2 = dx_2[4:N]
dx_3 = dx_3[4:N]

res1 = lm(Y~dY_1+x+cos1+sin1+cos2+sin2)  # 5 parms
res2 = lm(Y~dY_1+dY_2+dY_3+x+dx_1+dx_2+dx_3+cos1+sin1+cos2+sin2) # 10 parms
res3 = lm(Y~dY_1+x+dx_1+dx_2+cos1+cos2+sin2) # 6 parms
summary(res1) # Simple model
summary(res2) # Complex model
summary(res3) # Third model
```
Compare the complex and nested models using an F test.
```{r}
anova(res1, res2)
```

Compare the non nested model to the other two using BIC and AIC.
```{r}
AIC(res1, res2, res3)

BIC(res1, res2, res3)

```

<b>By looking at the F-test we see that we get a p-value < 0.01 significance level, which means we can reject the Null hypothesis and both models are not same. By looking at the AIC/BIC we can see that the third model is best as it has lowest AIC and BIC scores</b>

<h3>Examine the residuals of the model to see if they are white noise using autocovariance on the residuals and specral analysis of the residuals.</h3>
```{r}
dRes = res3$residuals
# hRes1 = residuals(res1)
plot.ts(dRes)

acf(dRes, 30, main="output")
Box.test(dRes,lag = 10, type = 'Box-Pierce')

spectrum(dRes, log="no")
```
 <b> By looking at the plots of the residuals it looks like the mean is 0 and variance is constant. The ACF and periodogram has no evidence of strong autocorrelation. The Box-Pierce test gives a p-value=0.054>0.05 significance, therefore we cannot reject the Null hypothesis that the autocorrelation at different lag components is 0. Thus we can conclude that the residuals are white noise</b>
 
Interpret the parameters of the model. What do they imply for the effect of the lagged feedback? Is the model stable? What do they imply for the effect of the input on the output?
```{r}
# To check for effect of input and lagged output on the model (i.e. dynamics) we will pass a step input 
# throught the difference equation model and observe the response. Since our model has sine, cosine pair we 
# expect some systematic oscillations in the output
N = 100 
x = array( 0, dim=c(N,1))
x[N/2] = 1 

# Create a step input
Sx = array(0, dim = c(N,1))
sum = 0
for (index in 1:N) {
  sum = sum + x[index]
  Sx[index] = sum 
}
x = Sx

Y = array( 0, dim=c(N,1))
YI = array( 0, dim=c(N,1))
# coefficients from our linear difference eq model
b0 = res3$coefficients[1]
b1 = res3$coefficients[2]
b2 = res3$coefficients[3]
b3 = res3$coefficients[4]
b4 = res3$coefficients[5]
b5 = res3$coefficients[6]
b6 = res3$coefficients[7]
b7 = res3$coefficients[8]

f1 = .1428571 # in previous assignment we found that f1 and f2
f2 = .2857143 # are two frequencies where the signal has external seasonal trend
time = 1:N
cos1 = cos(2*pi*time*f1)
sin1 = sin(2*pi*time*f1)
cos2 = cos(2*pi*time*f2)
sin2 = sin(2*pi*time*f2)

for (index in 1:N) {
  if (index == 1) {
    Y[index] = 1
    YI[index] = 1
    }
  if (index ==2 ){
    Y[index] = b0 + b1*Y[index-1]  + b2*x[index] + b3*x[index-1] + b5*cos1[index] + b6*cos2[index] + b7*sin2[index]
    YI[index] = b0 + b1*YI[index-1]  + b2*x[index] + b3*x[index-1]
    }

  if (index >2){
    Y[index] = b0 + b1*Y[index-1]  + b2*x[index] + b3*x[index-1] + b4*x[index-2] + b5*cos1[index] + b6*cos2[index] + b7*sin2[index]
    YI[index] = b0 + b1*YI[index-1]  + b2*x[index] + b3*x[index-1] + b4*x[index-2]
    }
}

layout( matrix( c(1,2), nrow=2, ncol=1, byrow="FALSE"))
plot(Y,type="l",xlab="time", main = 'Response of the model (in presence of sinusoids)')
plot(YI,type="l",xlab="time", main = 'Impulse response of the model (in absence of sinusoids)')
```
```{r}
F = c(res3$coefficients[2] ) # Since we have only 1 feedback term in our best model our
F = array(F, dim=c(1,1)) # feedback matrix is 1x1
ev<-eigen(F) # Stability of the model is determined by the lagged feedback terms
# ev$val
abs(ev$val)
 
```
<b> Since the absolute value of the eigen value of the feedback matrix is <1 the system is stable. This is also justified by the response of the model where the systematic oscillations are due to the sinusoids used to model external seasonal trends, but otherwise we can see that output oscillates about a equilibrium value. In absence of sinusoid inputs we can see the model decays to an equilibrium</b>

Plot the model predictions for the output and compare these to the observed.
```{r}
plot(demand[4:365], type = 'l', col = 'red', xlab = 'time', ylab = 'elec demand' )
lines(res3$fitted.values,col="green")
legend(300, 100, legend=c("Observed", "Fitted"),
       col=c("red", "green"), lty = 1:1, cex=0.8)

```